---
title: "Introduction to Business Analytics with R"
author: "Stephen Bernardo"
date: "2023-09-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1
Use a code chunk to create a datetime object d from the following character string that represents November 14, 2025: “11/20/2025”. Only use base R. Do not use functions from the lubridate package.

```{r}
d <- c("11/20/2025")
```

This code chunk will create the d dataframe

## Task 2
Use a new code chunk to check the data type of the d object by printing it. If it is in a date format, then it will show up as “2025-11-20”.

```{r}
class(d)
```

Checking what is the type of data for d

## Task 3
If necessary, install the lubridate package using the Package pane in RStudio. In a new code chunk, load the lubridate package.

```{r}
library(lubridate)
```

This is to load installed lubridate package

## Task 4
In a new code chunk, use the appropriate functions from the lubridate package to extract the year, month number, week number and weekday number from the object d created above. Save them as objects name d_year, d_month, d_week, and d_day respectively. Include code to display each of the four objects. 

```{r}
#Parsing the date first as it's in character format
d_parsed <- mdy(d)

#Extracting the year, month, week, date
d_year <- year(d_parsed)
d_month <- month(d_parsed)
d_week <- week(d_parsed)
d_day <- day(d_parsed)
```

This is to create new object as per instruction

## Task 5
Use a new code chunk to create another datetime object, d_25, using the object d, where d_25 is the date 25 days from now. Include code to display d_25.

```{r}
d_25 <- mdy(d) +25
print(d_25)
```

This is to show the date that is 25 days after d

## Task 6
Finally, use a new code chunk to calculate and display the difference between d and d_25 using the difftime function and check whether the difference is 25 days.

```{r}
print(difftime(d_25,mdy(d)))
```

To confirm that the difference between d_25 and d is 25 days.