---
title: "Twitter Sentiment Analysis"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
---

## ðŸ§­ Overview & Setup

This notebook walks you through on:

1. How to run a sentiment analysis on tweets using the VADER package
2. How to interpret the VADER scores
3. How to view top positive/negative tweets
4. How to visualize the compound scores distribution



### ðŸ“¦ Install and load packages

Install `tidyverse` and `vader` if you do not have them in your R environment.

- `tidyverse` is a collection of R packages for data science
- `vader` (Valence Aware Dictionary and sEntiment Reasoner) is a rule-based sentiment analysis tool specifically attuned to social media text

```{r message=FALSE, warning=FALSE}
# uncomment and run the lines below if you need to install these packages
# install.packages("tidyverse")
# install.packages("vader")
```

Load packages.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(vader)
```



### ðŸ“ƒ Read CSV file

```{r}
df_tweets = read_csv('lululemon-tweets.csv')
df_tweets %>% head()
```

Print out the number of rows.

```{r}
nrow(df_tweets)
```



## ðŸ”® Sentiment Analysis

### ðŸ“Œ Sample usage

To analyze a piece of text using VADER, use the `get_vader()` function. Here is an example using one of the tweets.

- `compound` score is the "overall" score between -1 (most extreme negative) and +1 (most extreme positive).
- `pos`, `neg`, and `neu` are ratios for proportions of text that fall in each category. These should all add up to be 1.
- We are mainly interested in the compound score.

```{r}
get_vader("Got my BeautyBase merch I am happy I got the first batch ðŸ¥°")
```



### ðŸ§® Calculate scores for all tweets

Because we have a DataFrame (`tibble`) that contains tweet texts in the `text` column, we will need to run `get_vader()` on each row of the `text` column. We can do this by using the `lapply()` function.

The code below stores the VADER results to a list variable named `vscores`. Note that this may take a while (a few minutes) if you have thousands of tweets.

```{r}
vscores <- df_tweets$text %>% lapply(get_vader)
```


Extract compound score and positive/neutral/negative percentages into separate columns.

```{r}
df_tweets <- df_tweets %>% mutate(
  compound = vscores %>% sapply(function(v) { as.numeric(v["compound"]) }),
  pos = vscores %>% sapply(function(v) { as.numeric(v["pos"]) }),
  neu = vscores %>% sapply(function(v) { as.numeric(v["neu"]) }),
  neg = vscores %>% sapply(function(v) { as.numeric(v["neg"]) }),
)
df_tweets
```



### ðŸ‘ 50 most positive tweets

```{r}
df_tweets %>% 
  arrange(desc(compound)) %>% 
  select(text, username, compound, pos, neu, neg) %>%
  head(50)
```



### ðŸ‘Ž 50 most negative tweets

```{r}
df_tweets %>% 
  arrange(compound) %>% 
  select(text, username, compound, pos, neu, neg) %>%
  head(50)
```



### âš–ï¸ Average compound score

```{r}
mean(df_tweets$compound)
mean(df_tweets$pos)
mean(df_tweets$neu)
mean(df_tweets$neg)
```



### ðŸ“¦ Box plot of compound scores

The box plot below gives you the spread and skewness of compound scores.

```{r}
df_tweets %>% 
  ggplot(aes(x=compound)) + 
  theme_classic() +
  geom_boxplot()
```



### ðŸ“Š Histogram of compound scores

The histogram below gives you an approximate distribution of the compound scores. 

```{r}
df_tweets %>% 
  ggplot(aes(x=compound)) + 
  theme_classic() +
  geom_histogram(bins = 20, color="black", fill="white")
```




